---
title: "Random Forest Model"
author: "Group F"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#Load Libraries
library(dplyr)
library(rpart)
library(rpart.plot)
library(caret)
library(caTools)
library(randomForest)
```

# Load the dataset

```{r}
# Load the dataset into the object "custData"
custData <- read.csv("CustData2_Prepared.csv")
custData$Eligible <- as.factor(custData$Eligible)
```

# Split the dataset into training data and testing data

```{r}
# Split the dataset to 80% training data and 20% testing data
set.seed(123)
train_index <- createDataPartition(custData$Eligible, p = 0.8, list = FALSE)
train_data <- custData[train_index, ]
test_data <- custData[-train_index, ]
```

# Train random forest model

```{r}
randomForest_model <- randomForest(Eligible ~ . -Annual_Salary,
                                   data = train_data, ntree = 100,
                                   mtry = 3, importance = TRUE)
```

# Visualize attribute importance

```{r}
#The attribute importance can be visualised using the random forest model
varImpPlot(randomForest_model)
```

# Make predictions using random forest

```{r}
randomForest_predictions <- predict(randomForest_model, newdata = test_data)
```

# Confusion matrix

## Create confusion matrix

```{r}
# Confusion Matrix for Random Forest
randomForest_cm <- confusionMatrix(as.factor(randomForest_predictions),
                                   as.factor(test_data$Eligible))
randomForest_matrix <- randomForest_cm$table
```

## Extract TP, TN, FP and FN from confusion matrix

```{r}
# Extract TruePositive, TrueNegative, FalsePositive 
# and FalseNegative for confusion matrix
randomForest_truePositive <- randomForest_matrix[1, 1]
randomForest_trueNegative <- randomForest_matrix[2, 2]
randomForest_falsePositive <- randomForest_matrix[1, 2]
randomForest_falseNegative <- randomForest_matrix[2, 1]
```

# Evaluation metrics

## Calculate the evaluation metrics

```{r}
# Calculate Evaluation Metrics
randomForest_accuracy <- 
  round(((sum(diag(randomForest_matrix)) / sum(randomForest_matrix))) * 100, 2)
randomForest_precision <- 
  round((randomForest_truePositive / (randomForest_truePositive +
                                        randomForest_falsePositive)) * 100, 2)
randomForest_recall <- 
  round((randomForest_truePositive / (randomForest_truePositive +
                                        randomForest_falseNegative)) * 100, 2)
randomForest_f1_score <- 
  round(2 * (randomForest_precision * randomForest_recall) /
                            (randomForest_precision + randomForest_recall), 2)
```

## Display the evaluation metrics

```{r}
cat("Random Forest Accuracy:", randomForest_accuracy, "% \n")
cat("Random Forest Precision:", randomForest_precision, "% \n")
cat("Random Forest Recall:", randomForest_recall, "% \n")
cat("Random Forest F1-score:", randomForest_f1_score, "% \n")
```

## Calculate Eligibility Rates - Random Forest

```{r}
# Assuming `predictions` is a vector of 1s (eligible) and 0s (not eligible) from your model
# For example: predictions <- predict(model, newdata, type = "response") > 0.5
 
# Count eligible customers
num_eligible_customers <- sum(randomForest_predictions == 1)

# Total number of customers
total_customers <- length(randomForest_predictions)
 
# Calculate the eligibility percentage
eligibility_percentage <- (num_eligible_customers / total_customers) * 100
 
cat("Percentage of eligible customers:", round(eligibility_percentage, 2), "%\n")
```

# Save the random forest model

```{r}
saveRDS(randomForest_model, file = "random_forest_model.rds")
```
